---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(magrittr)
```

# Datenimport

## Import CSV
Issue on line 5959, apprach import as lines and make modification via regex, 
then import as csv.

```{r}
lines <- readLines("./ugz_luftqualitaetsmessung_seit-2012.csv")

patn <- regex(",([0-9.]+)(\\d{4}-\\d{2}-\\d{2},.+)")
missing_newline_index <- grep(patn, lines)
lines[missing_newline_index] <- gsub(patn, ",\\1\n\\2", lines[missing_newline_index])

df <- read.csv(text=lines, na.strings = "NaN")
```

Input CSV has multiple headers per column, approach is to merge the location 
and typeinto a new header.

```{r}
columns <- names(df)
columns <- columns[-1]

names(df) <- c("datum", lapply(columns, function(col) paste(df[1,col], df[2,col], sep=".")))

df <- df[-c(1,2,3,4,5),]
```

## Convert from long to wide
Problem is that location is encoded in the column name.

```{r}
dfl <- df %>%
  pivot_longer(
    !datum,
    names_to = c("standort","name"),
    names_pattern = "(.+)\\.(.+)",
    values_to = "value",
    values_drop_na = FALSE
  )
```

Replace ugly german names

```{r}
dfl <- dfl %>%
  mutate(name = replace(name, name == "Schwefeldioxid", "so2")) %>%
  mutate(name = replace(name, name == "Kohlenmonoxid", "co")) %>%
  mutate(name = replace(name, name == "Ozon, höchstes Stundenmittel", "o3_max_h1")) %>%
  mutate(name = replace(name, name == "Ozon, Anzahl Stundenmittel > 120 µg/m3", "o3_nb_h1gt120")) %>%
  mutate(name = replace(name, name == "Stickstoffdioxid", "no2")) %>%
  mutate(name = replace(name, name == "Stickstoffmonoxid", "no")) %>%
  mutate(name = replace(name, name == "Feinstaub PM10", "pm10")) %>%
  mutate(name = replace(name, name == "Lufttemperatur", "t")) %>%
  mutate(name = replace(name, name == "Luftfeuchte relativ", "hr")) %>%
  mutate(name = replace(name, name == "Luftdruck", "p")) %>%
  mutate(name = replace(name, name == "Windgeschwindigkeit skalar", "wvs")) %>%
  mutate(name = replace(name, name == "Globalstrahlung", "strglo")) %>%
  mutate(name = replace(name, name == "Regendauer", "raindur"))
```

Normalize Standort
```{r}
dfl <- dfl %>%
  mutate(standort = str_replace(standort, "strasse", "")) %>%
  mutate(standort = str_replace(standort, "Zch_", "")) %>%
  mutate(standort = str_to_lower(standort))
```

Ensure data types
```{r}
dfl$datum %<>% as.Date
dfl$standort %<>% as.factor
dfl$value %<>% as.numeric
```

Convert back to wide
```{r}
dfw <- dfl %>%
  pivot_wider(names_from = name, values_from = value)
```

Validation
```{r}

```

# Erweiterung des Datensatzes
